Output Summary

### 2021-07-27 03-post-process/output-plots.Rmd
Update and add to output plots -- more info relative to fecundity inclusion

# Spawner-to-Parr Relationship

All red components are posterior summaries (medians and 95% credible intervals). The black curve is from an `nls()` fit on reconstructed parr and spawners -- the grey region is the confidence interval around this fit (from a parametric bootstrap).

```{r spawners-to-parr, fig.width = 6, fig.height = 4}
# extract posterior samples of the BH productivity and capacity parameters
bh_params = post_subset(post, c("alpha", "beta"), matrix = T)

@@

@@

# predict parr recruits along a sequence of spawner abundances
Sa_pred = seq(0, max(Sa_tot[5,]), length = 30)
Pb_pred = t(sapply(1:post_dim(post, "saved"), function(i) BH(Sa_pred, bh_params[i,1], bh_params[i,2])))
colnames(Pb_pred) = paste0("Pb_pred[", 1:length(Sa_pred), "]")
Pb_pred = post_convert(cbind(postpack:::id_mat(post), Pb_pred))
Pb_pred = post_summ(Pb_pred, "Pb_pred")

# fit ordinary least squares version
ols_fit = fit_basic_BH(jags_data)$BH_fit

@@

# calculate predicted curve at each pair/summarize
ols_R_pred = t(sapply(1:1000, function(i) BH(Sa_pred, ols_params[i,1], ols_params[i,2])))
ols_R_pred = apply(ols_R_pred, 2, function(x) quantile(x, c(0.5, 0.025, 0.975)))

@@

# empty plot
plot(1, 1, type = "n", xlim = range(0, Sa_tot[5,]), ylim = range(0, Pb[5,], Pb_pred[5,]),
     xlab = "Spawners", ylab = "Parr Recruits")

# draw fitted relationship: Bayesian
polygon(c(Sa_pred, rev(Sa_pred)), c(Pb_pred[4,], rev(Pb_pred[5,])), col = alpha("salmon", 0.5), border = NA)
lines(Pb_pred[4,] ~ Sa_pred, col = "red", lty = 2)
lines(Pb_pred[5,] ~ Sa_pred, col = "red", lty = 2)

# draw fitted relationship: OLS
polygon(c(Sa_pred, rev(Sa_pred)), c(ols_R_pred[2,], rev(ols_R_pred[3,])), col = alpha("grey", 0.5), border = NA)
lines(ols_R_pred[1,] ~ Sa_pred, lwd = 2)
lines(ols_R_pred[2,] ~ Sa_pred, lty = 2)
lines(ols_R_pred[3,] ~ Sa_pred, lty = 2)
lines(Pb_pred[3,] ~ Sa_pred, col = "red", lwd = 2)

# draw state pairs + uncertainty
segments(Sa_tot[4,], Pb[3,], Sa_tot[5,], Pb[3,], col = "red")
segments(Sa_tot[3,], Pb[4,], Sa_tot[3,], Pb[5,], col = "red")
points(Pb[3,] ~ Sa_tot[3,], pch = 21, col = "red", bg = alpha("salmon", 0.5))
```

@@

```{r sar-plot}
# function to calculate posterior 
get_SAR_post = function(post, jags_data) {
  # extract posterior abundances
  Ra_post = post_subset(post, "^Ra[", T) # adults returning to Columbia River mouth
  Ma_post = post_subset(post, "^Ma[", T) # smolts at top of LGD

@@

```{r}
par(oma = c(1,1,0,0))
plot_tseries(est = post_summ(post, "Pb_per_Sa_tot["), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Brood Year")
mtext(side = 2, outer = T, line = 0, "Parr/Spawner")
```

@@

```{r}
par(oma = c(1,1,0,0))
plot_tseries(est = post_summ(post, "Mb_per_Sa_tot["), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Brood Year")
mtext(side = 2, outer = T, line = 0, "Smolt/Spawner")
```

@@

```{r}
all_yrs = as.numeric(dimnames(jags_data$Pa_obs)[[1]])

@@

```{r stray-tot-plot}
par(oma = c(1,1,0,0))
plot_tseries(post_summ(post, "n_stray_tot"), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Return Year")
mtext(side = 2, outer = T, line = 0, "Number of Strays")
```

@@

```{r}
# extract posterior summaries
alpha_summ = post_summ(post, "alpha", digits = 0)
beta_summ = post_summ(post, "^beta$", digits = -3)
beta_per_peu_summ = post_summ(post, "beta_per_peu", digits = -1)
sigma_summ = post_summ(post, "sigma_Pb", digits = 2)

>>>>

# Spawner-to-Parr Relationship {.tabset .tabset-pills}

## Beverton-Holt

All red components are posterior summaries (medians and 95% credible intervals). The black curve is from an `nls()` fit on reconstructed parr and roughly approximated total egg production -- the grey region is the confidence interval around this fit.

```{r eggs-to-parr, fig.width = 6, fig.height = 4}
# extract posterior samples of the BH productivity and capacity parameters
bh_params = post_subset(post, c("alpha", "^beta$"), matrix = T)

@@

# extract posterior summaries of total egg production by brood year
f_tot = post_summ(post, "^f_tot[")

@@

# predict parr recruits along a sequence of spawner abundances
Sa_pred = seq(0, max(Sa_tot[5,]), length = 30)
f_pred = seq(0, max(f_tot[5,]), length = 30)

Pb_pred = t(sapply(1:post_dim(post, "saved"), function(i) BH(f_pred, bh_params[i,1], bh_params[i,2])))
colnames(Pb_pred) = paste0("Pb_pred[", 1:length(Sa_pred), "]")
Pb_pred = post_convert(cbind(postpack:::id_mat(post), Pb_pred))
Pb_pred = post_summ(Pb_pred, "Pb_pred")

# fit ordinary least squares version
ols_fit = fit_fecund_BH(jags_data)$BH_fit

@@

# calculate predicted curve at each pair/summarize
ols_R_pred = t(sapply(1:1000, function(i) BH(f_pred, ols_params[i,1], ols_params[i,2])))
ols_R_pred = apply(ols_R_pred, 2, function(x) quantile(x, c(0.5, 0.025, 0.975)))

@@

# empty plot
plot(1, 1, type = "n", xlim = range(0, f_tot[5,]), ylim = range(0, Pb[5,], Pb_pred[5,]),
     xlab = "Total Egg Production (millions)", ylab = "Parr Recruits", xaxt = "n")

x_ticks = axisTicks(par("usr")[1:2], log = F)
axis(side = 1, at = x_ticks, labels = x_ticks/1e6)

# draw fitted relationship: Bayesian
polygon(c(f_pred, rev(f_pred)), c(Pb_pred[4,], rev(Pb_pred[5,])), col = alpha("salmon", 0.5), border = NA)
lines(Pb_pred[4,] ~ f_pred, col = "red", lty = 2)
lines(Pb_pred[5,] ~ f_pred, col = "red", lty = 2)

# draw fitted relationship: OLS
polygon(c(f_pred, rev(f_pred)), c(ols_R_pred[2,], rev(ols_R_pred[3,])), col = alpha("grey", 0.5), border = NA)
lines(ols_R_pred[1,] ~ f_pred, lwd = 2)
lines(ols_R_pred[2,] ~ f_pred, lty = 2)
lines(ols_R_pred[3,] ~ f_pred, lty = 2)
lines(Pb_pred[3,] ~ f_pred, col = "red", lwd = 2)

# draw state pairs + uncertainty
segments(f_tot[4,], Pb[3,], f_tot[5,], Pb[3,], col = "red")
segments(f_tot[3,], Pb[4,], f_tot[3,], Pb[5,], col = "red")
points(Pb[3,] ~ f_tot[3,], pch = 21, col = "red", bg = alpha("salmon", 0.5))
```

## Total Eggs vs. Total Spawners

This figure illustrates that total spawners is a reasonable index of total egg production. The three lines illustrate the relationship if each spawner (regardless of age/sex) had a fixed number of eggs.

```{r, fig.width = 6, fig.height = 4}
# extract posterior summaries of total egg production by brood year
f_tot = post_summ(post, "^f_tot[")

par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(3,3,1,1))
plot(f_tot["mean",] ~ Sa_tot["mean",],
     ylim = c(0, max(f_tot["97.5%",])),
     xlim = c(0, max(Sa_tot["97.5%",])), type = "n",
     ylab = "Total Egg Production (Millions)", xlab = "Total Spawners", las = 1, yaxt = "n")
y_ticks = axisTicks(par("usr")[3:4], log = F)
axis(side = 2, at = y_ticks, labels = y_ticks/1e6, las = 2)

segments(Sa_tot["2.5%",], f_tot["mean",], Sa_tot["97.5%",], f_tot["mean",], col = alpha("red", 0.5))
segments(Sa_tot["mean",], f_tot["2.5%",], Sa_tot["mean",], f_tot["97.5%",], col = alpha("red", 0.5))
points(f_tot["mean",] ~ Sa_tot["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
abline(c(0, 1000), lty = 1)
abline(c(0, 1800), lty = 2)
abline(c(0, 3600), lty = 3)
legend("bottomright", title = "If Each Spawner Had:",
       legend = c("1000 Eggs", "1800 Eggs", "3600 Eggs"),
       lty = c(1,2,3), bty = "n")
```

## Parr per Egg

This figure shows the inherent density dependence built in to the Beverton-Holt recruitment function: with increasing egg density, survival per egg decreases. The horizontal line/region shows the theoretical maximum ($\alpha$) in the absence of density dependence.

```{r, fig.width = 6, fig.height = 4}
# extract posterior summaries of total egg production by brood year
f_tot = post_summ(post, "^f_tot[")

# conversion rates
Pb_per_f_tot = post_summ(post, "^Pb_per_f_tot[")

# max survival
alpha = post_summ(post, "^alpha")

par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(3,3,1,1))
plot(Pb_per_f_tot["mean",] ~ f_tot["mean",],
     ylim = c(0, max(Pb_per_f_tot["97.5%",], alpha["97.5%",])),
     xlim = c(0, max(f_tot["97.5%",])), type = "n",
     ylab = "Parr per Egg", xlab = "Total Egg Production (Millions)", las = 1, xaxt = "n")
x_ticks = axisTicks(par("usr")[1:2], log = F)
axis(side = 1, at = x_ticks, labels = x_ticks/1e6)

segments(f_tot["2.5%",], Pb_per_f_tot["mean",], f_tot["97.5%",], Pb_per_f_tot["mean",], col = alpha("red", 0.5))
segments(f_tot["mean",], Pb_per_f_tot["2.5%",], f_tot["mean",], Pb_per_f_tot["97.5%",], col = alpha("red", 0.5))
points(Pb_per_f_tot["mean",] ~ f_tot["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)

rect(par("usr")[1], alpha["2.5%",], par("usr")[2], alpha["97.5%",], border = alpha("red", 0.5), col = alpha("salmon", 0.5))
abline(h = alpha["mean",], lwd = 2, col = "red")
box()
```

This figure is identical to above, but shows parr production per spawner instead of per egg.

```{r, fig.width = 6, fig.height = 4}
# extract posterior summaries of total spawners by brood year
Sa_tot = post_summ(post, "^Sa_tot[")

# conversion rates
Pb_per_Sa_tot = post_summ(post, "^Pb_per_Sa_tot[")

par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(3,3,1,1))
plot(Pb_per_Sa_tot["mean",] ~ Sa_tot["mean",],
     ylim = c(0, max(Pb_per_Sa_tot["97.5%",])),
     xlim = c(0, max(Sa_tot["97.5%",])), type = "n",
     ylab = "Parr per Spawner", xlab = "Spawners", las = 1)
segments(Sa_tot["2.5%",], Pb_per_Sa_tot["mean",], Sa_tot["97.5%",], Pb_per_Sa_tot["mean",], col = alpha("red", 0.5))
segments(Sa_tot["mean",], Pb_per_Sa_tot["2.5%",], Sa_tot["mean",], Pb_per_Sa_tot["97.5%",], col = alpha("red", 0.5))
points(Pb_per_Sa_tot["mean",] ~ Sa_tot["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
```

@@

```{r sar-plot, fig.width = 6, fig.height = 4}
# function to calculate posterior 
get_SAR_post = function(post, jags_data) {
  # extract posterior abundances
  Ra_post = post_subset(post, "^Ra[", T) # adults returning to Columbia River mouth
  Ma_post = post_subset(post, "^Ma[", T) # smolts at top of LGD

@@

```{r, fig.width = 6, fig.height = 4}
par(oma = c(1,1,0,0))
plot_tseries(est = post_summ(post, "Pb_per_Sa_tot["), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Brood Year")
mtext(side = 2, outer = T, line = 0, "Parr/Spawner")
```

@@

```{r, fig.width = 6, fig.height = 4}
par(oma = c(1,1,0,0))
plot_tseries(est = post_summ(post, "Mb_per_Sa_tot["), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Brood Year")
mtext(side = 2, outer = T, line = 0, "Smolt/Spawner")
```

@@

```{r, fig.width = 6, fig.height = 4}
all_yrs = as.numeric(dimnames(jags_data$Pa_obs)[[1]])

@@

```{r stray-tot-plot, fig.width = 6, fig.height = 4}
par(oma = c(1,1,0,0))
plot_tseries(post_summ(post, "n_stray_tot"), yrs = as.numeric(rownames(jags_data$Pa_obs)[observable]))
mtext(side = 1, outer = T, line = -0.5, "Return Year")
mtext(side = 2, outer = T, line = 0, "Number of Strays")
```

@@

```{r}
# extract posterior summaries
alpha_summ = post_summ(post, "alpha", digits = 3)
beta_summ = post_summ(post, "^beta$", digits = -3)
beta_per_peu_summ = post_summ(post, "beta_per_peu", digits = -1)
sigma_summ = post_summ(post, "sigma_Pb", digits = 3)
